---
title: "Conjoint Analysis"
author: "Ashwin Malshe"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(radiant)
```



## Design of experiment

```{r}
all_profile <- doe(c("room; Small Suite; Large Room; Room Office",
                      "bus_amenities; Internet access; Speaker phone; Room fax",
                      "leisure; Exercise room; Pool; Exercise room and Pool",
                      "extras; Shoe shine; Tape library; Fruit and cheese; Newspaper",
                      "rest_delivery; No; Yes"), seed = 777)
```




```{r}
all_profile_dt <- summary(all_profile)
```


`Radiant` changed all the levels with spaces into underscores. This will create problems later on.

```{r}
all_profile_dt <- all_profile_dt %>% 
  mutate(across(everything(), ~stringr::str_replace_all(.x, "_", " ")))
```



Import the profiles

```{r}
forte_plan <- readxl::read_excel("Forte Hotel.xlsx", sheet = "Plan-extended") %>% 
  mutate(across(everything(), as.factor))
```







```{r}
forte <- readxl::read_excel("Forte Hotel.xlsx", sheet = "Preference") %>% 
  mutate(repondent = row_number()) %>% 
  tidyr::pivot_longer(starts_with("Bundle"), names_to = "profile", values_to = "preference") %>% 
  mutate(profile = as.factor(profile))
```



```{r}
forte_full <- inner_join(forte, forte_plan, by = "profile")
```


## Conjoint analysis

```{r}
conjoint1 <- forte_full %>% 
  conjoint(rvar = "preference", evar = c("room", "bus_amenities",	"leisure",	"extras",	"rest_delivery"))
```


```{r}
summary(conjoint1)
```

Plot the results

```{r}
plot(conjoint1)
```





```{r}
predict(conjoint1, all_profile_dt) %>% 
  arrange(-Prediction)
```


## Business simulations

Although from all the possible profiles we can get the most valued profile, it may not be practical or profitable. The marketing managers may have a few business profiles identified before the analysis.

Let's figure out which one of these is the best profile.

```{r}
forte_sim <- readxl::read_excel("Forte Hotel.xlsx", sheet = "Simulation Profiles") %>% 
  mutate(across(everything(), as.factor))
```



```{r}
forte_sim$Simulations %>% 
  cbind(predict(conjoint1, forte_sim) )

```
