---
title: "SFA"
author: "Ashwin Malshe"
date: "1/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(frontier)
```


```{r}
mex_dt <- haven::read_dta("mexico-cstat.zip") %>% 
  mutate(cyear = lubridate::year(datadate))
```

```{r}
count(mex_dt, cyear)
```

```{r}
summary(mex_dt$sale)
```

We will use a subset of this data set. The output variable is `sale` and the inputs are the assets (`at`), selling and general administrative expenses (`xsga`), which some people use as a proxy for marketing, and number of employees (`emp`).

Due to limited number of observations, we will combine observations across the years. This is not technically right and we should use a panel model for this.

```{r}

mex_dt2 <- mex_dt %>% 
  select(sale, at, xsga, emp, sic) %>% 
  mutate(across(everything(), ~ ifelse(.x <= 0, NA, .x))) %>% # Set negative or zero values to NA
  tidyr::drop_na()  %>% # Drop missing observations.
  mutate(onedsic = factor(as.integer(sic) %/% 1000)) # We will use it in the second SFA model
```


```{r}
lm(sale ~ at + xsga + emp, data = mex_dt2) %>%  summary()
```

## SFA

For running an SFA model, we have to use logged values for all the variables.

```{r}
model_sfa <- sfa(log(sale) ~ log(at) + log(xsga) + log(emp), data = mex_dt2)
```

```{r}
summary(model_sfa)
```


## Obtain technical efficiency estimates

```{r}
model_eff <- efficiencies(model_sfa)
```

```{r}
head(model_eff)
```


Plot these efficiencies as histograms

```{r}

model_eff %>% 
  as.data.frame() %>% 
  ggplot(aes(efficiency)) +
  geom_histogram(color = "white") +
  theme_minimal()

```


## Modeling the determinants of efficiency

As a variation of the basic SFA model, we can make the mean of the inefficiency term a function of some other variables. For instance, if we believe that the inefficiency is a function of the industry in which companies operate, we can model that directly. 

In the following code, we will use `onedsic`

```{r}
model_sfa2 <- sfa(log(sale) ~ log(at) + log(xsga) + log(emp) | onedsic, data = mex_dt2)
```

```{r}
summary(model_sfa2)
```


```{r}
model_eff2 <- efficiencies(model_sfa2, margEff = TRUE)
```

```{r}
head(model_eff2)
```

How do we extract the marginal effects? They are stored in `model_eff2` as `attributes`

```{r}
model_eff2_me <- attributes(model_eff2)
```

```{r}
head(model_eff2_me$margEff)
```

```{r}
model_eff2_me$margEff %>% 
  as.data.frame() %>% 
  summary()
```


```{r}
model_sfa3 <- sfa(log(sale) ~ log(at) + log(xsga) + log(emp) | onedsic - 1, data = mex_dt2)
```

```{r}
summary(model_sfa3)
```


